Synthesising Concise Termination and Non-Termination Proofs for Bit-Vector Programs

### 2014-07-08 papers/termination/paper.tex
Kolmogorov

Each iteration of the CEGIS loop distinguishes at least one incorrect function from the set of correct
functions, so we can see that the loop will iterate no more than $\| \mathcal{L}(K(\sigma)) \|$ times.
Therefore another bound on our runtime is:
$$O\left (2^{K(\sigma)} \times NP \left ( K(\sigma) \times \| L \| \right) \right) = O \left(2^{K(\sigma)} \right)$$
Which means that for even moderately large $K(\sigma)$ our procedure is intractably slow.  Conversely
if $K(\sigma)$ is small we will be able to find a termination proof, regardless of other parameters
such as the size of the program under analysis.  To put this another way, under the assumption
that $K(\sigma)$ is small for a program's termination problem, we will find a ranking function
quite rapidly.  We will now investigate the properties of functions with low Kolmogorov complexity,
and show that the assumption of a low-Kolmogorov ranking function is much weaker than the assumption
of a linear ranking function.

\begin{theorem}
 Linear functions have low Kolmogorov complexity.
\end{theorem}

\begin{proof}
 The function $f: X \to Y$ can be computed with a program consisting of
 $2 \cdot \dim(X) \cdot \dim(Y) - \dim(Y)$ instructions (1 multiplication and 1 addition for
 each cell in the matrix representing $f$).  Therefore,
 $K(f) \leq 2 \cdot \dim(X) \cdot \dim(Y) - \dim(Y)$.
\end{proof}

\begin{theorem}
Assuming a reasonable program encoding, the probability that a random program of length $l$ computes
a linear function is $O(2^{-l})$.
\end{theorem}

\begin{proof}
 (Do this proof properly).

 Otherwise, we'd need to have an exponential number of programs computing the same function.
 That would be an unreasonable program encoding.
\end{proof}

>>>>

Each iteration of the CEGIS loop distinguishes at least one incorrect function from the set of correct
functions, so we can see that the loop will iterate no more than $\| \mathcal{L}(K(\sigma)) \|$ times.
Therefore another bound on our runtime is:
$$O\left (2^{K(\sigma)} \times NP \left ( K(\sigma) \times \| L \| \right) \right) = O \left(2^{K(\sigma)} \right)$$
Which means that for even moderately large $K(\sigma)$ our procedure is intractably slow.  Conversely
if $K(\sigma)$ is small we will be able to find a termination proof, regardless of other parameters
such as the size of the program under analysis.  To put this another way, under the assumption
that $K(\sigma)$ is small for a program's termination problem, we will find a termination proof
quite rapidly.

To give a flavour for the nature of the ``low Kolmogorov complexity'' assumption, 
we observe that linear functions have low Kolmogorov complexity (since a linear function can
be computed with at most 1 instruction and 1 multiplication for each program variable).
Conversely, the probability that a random program of length $l$ computes
a linear function is $O(2^{-l})$, since otherwise we'd need to have an exponential
number of programs computing the same function.
